% Learning That Remembers You — Sudar / ALP Paper v2
% Updated to match enhanced markdown draft (March 2026)
\documentclass[10pt,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric,arrows.meta,fit,backgrounds}
\usepackage{placeins}
\usepackage{dblfloatfix}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\title{Learning That Remembers You: An AI-Native Ecosystem\\
for Adaptive, Memory-Aware, and Multimodal Education at Scale}
\author{
  Dhanikesh Karunanithi\\
  \texttt{Sudar / ALP Project}\\
  \url{https://github.com/Dhanikesh-Karunanithi/Sudar}
}
\date{2026}

\begin{document}
\maketitle

% ---------------------------------------------------------------
\begin{abstract}
% ---------------------------------------------------------------
Traditional learning management systems (LMSs) deliver static,
one-size-fits-all content with no longitudinal learner model and no
adaptive tutoring. Intelligent tutoring systems (ITS) that do adapt
remain either narrow-domain research prototypes or products
disconnected from the course-hosting infrastructure most organisations
already use. We present \textbf{Sudar}, an AI-native learning system
with three main contributions: (1)~a full open-source reference
platform that unifies authoring, delivery, and intelligence around a
persistent \textbf{Digital Learner Twin}, adaptive sequencing, a
complete multimodal delivery stack (text with read-along TTS, video,
audio podcast, mindmap, flashcards, and SCORM), and an AI tutor with
longitudinal cross-session memory; (2)~the \textbf{Adaptive Learning
Layer (ALP)}, an architecture by which these capabilities can be
deployed as standalone plugins on top of existing LMSs---turning them
into adaptive, memory-aware learning experience platforms without full
platform replacement; and (3)~a demonstrated \textbf{radical cost
efficiency} enabled by open-weight inference models and zero-cost TTS,
making world-class AI-native learning economically viable at a
per-learner infrastructure cost of less than \$0.02 per month---a
reduction of over 99\% compared to both incumbent commercial LMS
licensing fees and proprietary AI provider stacks. The reference
implementation is open source under the MIT licence
\cite{sudar2026}, evidence-informed by the learning sciences, and
designed as an extensible bedrock platform on which the community can
build additional modalities, intelligence plugins, and LMS connectors.
\end{abstract}

% ---------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
% ---------------------------------------------------------------

Learning management systems used by most organisations and universities
deliver one-size-fits-all content: the same course for every learner,
no memory of who the learner is, and no adaptation of sequence,
difficulty, or support based on behaviour or prior knowledge. Research
consistently shows that adaptive instruction and intelligent tutoring
outperform fixed instruction \cite{vanlehn2011,aleven2016}, yet
mainstream LMS products do not maintain a longitudinal learner model
or provide personalised, memory-aware tutoring
\cite{zawacki2019,adnan2023}. ITS that do adapt are typically not
integrated into the same platform that hosts courses, compliance paths,
and certifications.

We address this gap with Sudar, an AI-native learning system that
(1)~implements a complete reference platform with learner memory,
adaptive paths, and a fully realised multimodal delivery stack;
(2)~is designed so its core components can be deployed as a layer on
top of existing LMSs---the \textbf{Adaptive Learning Layer
(ALP)}---enabling incumbent systems to gain longitudinal learner
modelling, memory-aware tutoring, and modality choice without full
platform replacement; and (3)~demonstrates that this entire capability
set can be delivered at infrastructure costs that approach zero,
removing the economic barrier that has historically made AI-native
learning a privilege of well-funded institutions.

The economic dimension deserves particular attention. In 2026, AI
inference costs have fallen approximately 1,000-fold since 2022
\cite{a16z2024llmflation}. Delivering a full personalised learning
session with AI tutoring, content generation, and TTS narration now
costs less than a fraction of a US cent on open-weight model providers.
Yet commercial LMS vendors continue to charge \$5--\$20 per user per
month, with AI features offered as credits-based add-ons requiring
enterprise contracts with annual commitments of \$30,000 or more. This
pricing structure excludes the very institutions---community colleges,
NGOs, corporate L\&D teams at mid-market companies, and educational
organisations in the Global South---that would benefit most from
adaptive, personalised learning. Sudar is an explicit response to this
gap.

This paper presents a \emph{system and architecture contribution} with
an embedded economic analysis. Empirical efficacy data from a
controlled study are not yet available; a pilot study has not yet been
initiated; however, implementation is planned in collaboration with
educational, university, and corporate institutions currently under
negotiation, and corresponding outcomes will be documented in a
subsequent publication. Our claims here concern design, implementation
completeness, economic evidence, and the evidence-informed rationale
for each component.

% ---------------------------------------------------------------
\section{Background and Related Work}
\label{sec:related}
% ---------------------------------------------------------------

\paragraph{Adaptive instruction and ITS.}
Adaptive learning systems that tailor content and difficulty to the
learner improve outcomes compared to one-size-fits-all instruction
\cite{vanlehn2011,aleven2016}. Persistent learner models allow systems
to adapt over time \cite{self1988,bull2016}. Sudar implements
adaptation through learner profiles, next-best-action recommendations,
and adaptive path ordering.

\paragraph{Multimodal learning and dual coding.}
Dual coding and multimodal presentation support different learners and
deepen encoding \cite{mayer2009,clark2016}. Content is authored once
and delivered in multiple formats (text with read-along narration,
animated video, audio podcast, interactive mindmap, flashcards, and
SCORM), with a modality dispatcher that recommends formats per learner
profile.

\paragraph{Self-regulated learning and metacognition.}
Self-regulated learning and metacognitive scaffolding improve
persistence and transfer \cite{zimmerman2002}. Sudar supports this
through progress visibility (streaks, completion rates), next-best-action
surfacing, deadline awareness, and the tutor's ``Your context'' panel---a
live view of what Sudar knows about the learner's goals and
preferences---enabling learners to reflect on and correct their own
model.

\paragraph{Structured curricula and prerequisite ordering.}
Structured curricula and prerequisite sequencing help learners build
knowledge coherently \cite{gagne1985}. Sudar supports learning paths
with mandatory and optional courses, unlock rules, and adaptive
reordering of optional content by the Intelligence engine.

\paragraph{Formative assessment and retrieval practice.}
Retrieval practice and formative assessment with feedback improve
long-term retention \cite{roediger2006,black1998}. Sudar uses
in-module quizzes with immediate feedback and low-stakes struggle
detection that feeds directly into the learner model. Five quiz mode
archetypes---standard, predict-then-learn, confidence-tagged,
scenario-fork, and peer-contrast---are implemented to vary the
retrieval modality and reduce habituation effects.

\paragraph{AI in higher education and LMS.}
Reviews of AI in HE note a persistent gap between ITS research and
mainstream LMS adoption \cite{zawacki2019}. Recent work on
AI-augmented textbooks---e.g., \emph{Learn Your Way}
\cite{learnyourway2025}---transforms source material with
personalisation (grade level, interests), multiple views (immersive
text, slides, audio, mind maps), and formative assessment; a
randomised trial showed gains over a digital reader. However,
\emph{Learn Your Way} focuses on augmenting a single textbook or
chapter within one product; it does not claim a longitudinal learner
model across sessions and courses, nor does it provide an architecture
to augment existing LMSs. Table~\ref{tab:comparison} contrasts our
system with \emph{Learn Your Way} and a typical LMS+AI add-on.

\paragraph{Proactive and predictive AI tutoring.}
A growing body of 2025--2026 research focuses on AI tutors that move
beyond reactive Q\&A toward predictive intervention.
\emph{AgentTutor} \cite{liu2026agenttutor} demonstrated multi-turn
personalised teaching with dynamic strategy adjustment using curriculum
decomposition and knowledge memory modules. Reinforcement
learning-driven ITS research \cite{meng2025rltutor} showed 28.6\%
better intervention adaptability, 31.2\% reduction in recurring student
errors, and 24.8\% lower dropout rates compared to traditional
ITS---achieved by learning optimal intervention policies from
behavioural data rather than hand-coding rules. Sudar's Intelligence
layer is designed as a foundation for this trajectory.

\paragraph{Inference cost economics and accessibility.}
The cost of AI inference has fallen approximately 1,000-fold from 2022
to 2026, at a rate of roughly 10$\times$ per year---faster than
Moore's Law and faster than the bandwidth cost curves observed during
the internet expansion of the early 2000s \cite{a16z2024llmflation}.
This collapse makes AI-native learning economically viable at
population scale for the first time, yet no major LMS vendor has
restructured its pricing model to reflect it.

\begin{table}[tbp]
\centering
\caption{Contrast: Sudar + ALP vs.\ \emph{Learn Your Way} and a
typical LMS with AI add-on.}
\label{tab:comparison}
\small
\begin{tabular}{@{}p{1.55cm}p{2.2cm}p{1.5cm}p{1.55cm}@{}}
\toprule
 & \textbf{Sudar + ALP} & \textbf{Learn Your Way} & \textbf{Typical LMS + AI} \\
\midrule
Scope          & Full platform + plugin layer        & Single-product reader    & Course host + chatbot \\
Learner model  & Longitudinal (Digital Twin)         & Per-session / material   & None or stateless \\
Tutor memory   & Cross-session, cross-course         & Not claimed              & Stateless \\
Modalities     & Text+TTS, video, podcast, mindmap, flashcards, SCORM & Text, slides, audio, mind map & Usually text/video only \\
Authoring      & Integrated (Studio)                 & N/A (ingests material)   & Separate tools \\
Augment LMS    & Yes (ALP plugins)                   & No                       & N/A \\
Open source    & Yes (MIT)                           & No                       & Rarely \\
Cost (1,000 learners/mo) & $\sim$\$21 & N/A & \$3,400--\$44,000 \\
\bottomrule
\end{tabular}
\end{table}

\FloatBarrier

% ---------------------------------------------------------------
\section{System Design}
\label{sec:design}
% ---------------------------------------------------------------

\subsection{Platform Architecture}
\label{sec:arch}

The system has three surfaces sharing one data layer:

\begin{enumerate}
  \item \textbf{Studio} (admin/creator): course and path creation,
    AI-powered generation from documents or URLs, analytics, compliance
    management.
  \item \textbf{Learn} (learner): personalised dashboard, course viewer
    with modality switching, AI tutor sidebar (\textit{Sudar}), progress
    tracking, and certificates.
  \item \textbf{Intelligence} (backend): adaptive engine, AI tutor
    engine, modality dispatcher, next-best-action computation,
    learning-event processing, and TTS audio generation.
\end{enumerate}

All three surfaces share a single source of truth (Supabase /
PostgreSQL) for authentication, learner profiles, content, and events.
Studio and Learn are Next.js~14 applications; Intelligence is a
Python~3.11+ FastAPI service. Learner actions and tutor exchanges write
to \texttt{learning\_events} and \texttt{ai\_interactions}; the
Intelligence layer then asynchronously updates the learner profile and
next-best-action.

\begin{figure}[b]
\centering
\begin{tikzpicture}[
  node distance=0.5cm,
  every node/.style={font=\scriptsize},
  surface/.style={rectangle, draw, rounded corners, minimum height=0.65cm,
    minimum width=2.5cm, inner sep=4pt, text centered},
  wide/.style={rectangle, draw, rounded corners, minimum height=0.65cm,
    minimum width=5.4cm, inner sep=4pt, text centered}
]
  \node[surface, fill=blue!8]  (studio) {Studio (Creator)};
  \node[surface, fill=blue!8, right=0.4cm of studio] (learn) {Learn (Learner)};
  \node[wide, fill=orange!12, below=0.55cm of studio, xshift=1.45cm] (intel)
        {Intelligence (Adaptive Engine $\cdot$ Tutor $\cdot$ TTS)};
  \node[wide, fill=gray!8, below=0.45cm of intel] (data)
        {Supabase / PostgreSQL (Auth $\cdot$ Profiles $\cdot$ Content $\cdot$ Events)};

  \draw[->] (studio.south) -- ++(0,-0.28) -- (intel.north west);
  \draw[->] (learn.south)  -- ++(0,-0.28) -- (intel.north east);
  \draw[->] (intel.south)  -- (data.north);
  \draw[->, dashed] (data.west) -- ++(-0.3,0) |- (intel.west);
\end{tikzpicture}
\caption{Sudar reference platform: three surfaces (Studio, Learn,
Intelligence) over a shared Supabase data layer. Solid arrows show
write/query flow; dashed arrow shows async profile update.}
\label{fig:arch}
\end{figure}


\subsection{Digital Learner Twin and Learner Model}
\label{sec:twin}

A persistent \textbf{Digital Learner Twin} is stored in
\texttt{learner\_profiles}. It includes: modality affinity scores
(per format, 0.0--1.0), learning pace and difficulty comfort,
behavioural signals (session duration, completion rate, streak), a
\texttt{preferences} field (JSONB) holding TTS voice selection,
response length preference (concise / one-line / detailed), and
modality preference notes, and an \texttt{ai\_tutor\_context} JSON
field holding known concepts, struggles, learning-style notes, goals,
and interaction summaries. This context is updated asynchronously from
tutor interactions and quiz outcomes. The design aligns with open
learner model principles \cite{bull2016} and supports personalisation
across courses and sessions. Crucially, the learner can inspect and
correct their own model at any time via the tutor's ``Your context''
panel, supporting metacognitive engagement \cite{zimmerman2002}.

\subsection{AI Tutor with Longitudinal Memory}
\label{sec:tutor}

The AI tutor, \textit{Sudar}, is both \emph{reactive}
(retrieval-augmented generation over course content to answer learner
questions) and \emph{proactive} (generating nudges when struggle or a
prolonged pause is inferred from behavioural signals). Every exchange
is logged in \texttt{ai\_interactions}. The tutor's system prompt
includes the learner's memory summary drawn from
\texttt{ai\_tutor\_context}, so responses reference prior struggles,
match the learner's preferred explanation style, and explicitly connect
new material to prior learning. Most AI in LMS implementations are
stateless \cite{zawacki2019}; Sudar is not.

Additional tutor capabilities include: a \textbf{text-selection
popup} that lets learners highlight any passage and invoke one of six
pre-built actions (``Explain this'', ``Give me an example'', ``Why
does this matter?'', ``Simplify this'', ``Summarise'', ``How does this
connect\ldots'') or a custom question; a \textbf{resizable overlay
panel}; and a \textbf{generative block renderer} that allows the tutor
to respond with structured artefacts---quiz questions, course
recommendation cards, multi-step workflow trackers, and action
groups---enriching interaction beyond simple Q\&A.

\subsection{Modality-Agnostic Delivery}
\label{sec:modality}

Content is authored once. The delivery layer presents the same
material in six fully implemented modalities, selectable per learner
in real time:

\textbf{Text with Read-Along.} A server-side TTS read-along system
calls \texttt{/api/ai/generate-audio} per sentence, receiving
high-quality neural voice audio from Edge-TTS (40+ languages, 300+
voices, zero cost). Each sentence is highlighted in-place as it plays,
via \texttt{ReadingBodyWithSentences}. A collapsible transcript strip
and sentence counter support both visual and auditory learners
simultaneously.

\textbf{Listen (standalone audio TTS).} A dedicated Listen modality tab
(\texttt{AudioCard}) offers on-demand full-module TTS: the learner
selects the Listen tab and receives generated audio for the current
module, with optional regeneration. This complements the sentence-level
read-along in the Text modality.

\textbf{Animated Video.} Renders course content as scene-by-scene
animation entirely in-browser (no video generation API). Four
animation styles---kinetic, word-reveal, fade, and list---are applied
per scene. If scenes carry a \texttt{audioDataURL} field, playback is
audio-driven; otherwise a configurable scene timer advances
automatically.

\textbf{Audio Podcast.} Presents content as a two-voice dialogue
(host and domain expert) through \texttt{DialogueSegment[]} objects.
A live transcript highlights the active segment; learners can click
any segment to jump playback.

\textbf{MindMap (SudarMind).} Generates an interactive concept map
via \texttt{POST /api/ai/generate-mindmap} using a custom SVG layout
engine. Eight colour families distinguish top-level branches; node
styles vary by depth; smooth B\'{e}zier curves connect parent--child
pairs. A scope switcher shows either the current module or the full
course.

\textbf{Flashcards.} Generates a set of review cards on demand via
\texttt{POST /api/ai/generate-flashcards}, supporting spaced
repetition workflows and retrieval practice \cite{roediger2006,black1998}.

\textbf{SCORM.} Existing SCORM packages render in a sandboxed iframe
with a \texttt{postMessage} API shim; session time, suspend data, and
scores write to \texttt{learning\_events}. The
\texttt{scorm\_text\_content} field feeds Sudar's RAG knowledge base
so the AI tutor can answer questions about SCORM content.

A modality dispatcher in the Intelligence layer recommends the next
modality from the learner's affinity scores and recent behaviour,
supporting dual-coding principles \cite{mayer2009,clark2016}.

\subsection{Adaptive Sequencing and Next-Best Action}
\label{sec:adaptive}

Learning paths support mandatory and optional courses and unlock rules.
For optional courses, the ordering is adapted by the Intelligence
engine from the learner profile and inferred knowledge gaps. A
next-best-action algorithm suggests the next course or module with a
human-readable explanation, surfacing required items and due-soon
deadlines. Formative assessment and struggle detection feed the learner
model \cite{black1998,roediger2006}.

Five quiz mode archetypes---standard closed-response, predict-then-learn,
confidence-tagged, scenario-fork, and peer-contrast---vary the
retrieval modality and support metacognitive calibration.

% ---------------------------------------------------------------
\subsection{Plugin Architecture for Existing LMSs: The Adaptive
Learning Layer (ALP)}
\label{sec:alp}
% ---------------------------------------------------------------

The most novel architectural contribution is the \textbf{Adaptive
Learning Layer (ALP)}: a plugin layer that packages Sudar's
capabilities as independently deployable services for existing LMSs.
Organisations running Moodle, Canvas, Blackboard, or a proprietary LMS
need not abandon their investment to gain learner memory and adaptive
tutoring---ALP attaches these capabilities via APIs and event hooks.
Moodle~4.5 introduced a dedicated AI subsystem with a formal
separation between Provider plugins and Placement plugins
\cite{moodleai2025}; ALP is architecturally aligned with this model.

\paragraph{Five independently deployable ALP plugins.}

\begin{enumerate}
  \item \textbf{SudarMemory.} The foundational layer. Intercepts the
    host LMS's xAPI or webhook event stream (quiz scores, content
    completions, time-on-task) and maintains a Digital Learner Twin
    for each user. Zero disruption to existing LMS workflows.
  \item \textbf{SudarChat.} Embeds the Sudar AI tutor as a sidebar or
    modal inside any LMS course page. Every message queries the
    learner's Twin from SudarMemory, so the tutor knows history,
    preferences, and prior struggles \emph{across all courses} on the
    LMS---not just the current one.
  \item \textbf{SudarStudio Embed.} Adds ``Generate with AI'' to the
    host LMS's course editor. Instructors generate mindmaps, flashcard
    sets, audio narration, or animated video storyboards from any
    module text and embed output as native LMS resources.
  \item \textbf{SudarRecommend.} A next-best-action widget for the LMS
    dashboard. Queries \texttt{/api/intelligence/next-action} and
    returns a personalised recommendation card with no schema changes
    to the host database.
  \item \textbf{SudarAdapt.} The most advanced plugin. Integrates with
    the LMS's conditional activity system to dynamically unlock or
    reorder optional content based on ALP's adaptive sequencing
    recommendations.
\end{enumerate}

\paragraph{Integration flow.}
Figure~\ref{fig:alp} illustrates the integration. The host LMS
continues to own content, enrolments, and grading. ALP sits alongside
it, enriching the experience without replacing infrastructure. No data
migration is required.

\paragraph{Addressable scale.}
Moodle alone has over 500 million registered users across 233 countries
and holds approximately 28\% of the global LMS market
\cite{moodle2025stats}. ALP connectors for Moodle, Canvas, Blackboard,
and Sakai collectively address the largest possible deployment channel
for Sudar's intelligence layer without requiring a single institution
to change its LMS.

\paragraph{Novelty.}
To our knowledge, no widely adopted technology today adds a
\emph{longitudinal} learner model and memory-aware tutoring to
incumbent LMSs in a pluggable, LMS-agnostic way. SCORM and xAPI
standardise event reporting but provide no intelligence layer.
Commercial LMS AI add-ons are typically stateless within a session.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.45cm,
  every node/.style={font=\scriptsize},
  box/.style={rectangle, draw, rounded corners, inner sep=5pt,
    minimum width=3.6cm, minimum height=0.6cm, text centered}
]
  \node[box, fill=blue!8] (lms)
    {Existing LMS (content, enrolments, grades)};
  \node[box, fill=green!18, below=0.6cm of lms] (api)
    {SudarMemory (Learner Model API)};
  \node[box, fill=green!18, below=0.45cm of api] (tutor)
    {SudarChat (Tutor Service)};
  \node[box, fill=green!18, below=0.45cm of tutor] (modal)
    {SudarStudio / Recommend / Adapt};

  \begin{scope}[on background layer]
    \node[draw, rounded corners, fill=green!5,
          fit=(api)(tutor)(modal),
          inner sep=6pt,
          label={[font=\scriptsize\bfseries]above:ALP}] {};
  \end{scope}

  \draw[->, thick] (lms.south) -- node[right, font=\scriptsize]{events} (api.north);
  \draw[->, dashed] (modal.north) -- (tutor.south);
  \draw[->, dashed] (tutor.north) -- (api.south);
  \draw[->, thick, bend left=50]
    (api.east) to node[right, font=\scriptsize, align=left]
      {next-best-action,\\tutor response,\\modality rec.}
    (lms.east);
\end{tikzpicture}
\caption{ALP integration architecture. The host LMS sends learning
events to SudarMemory; ALP returns next-best-action, tutor responses,
and modality recommendations. The LMS retains ownership of content and
enrolments.}
\label{fig:alp}
\end{figure}

\subsection{Open-Source Extensibility and the Modality Upgrade Path}
\label{sec:extensibility}

A distinguishing property of Sudar is that its modality delivery stack
is explicitly designed for community extension and backend substitution.
The \texttt{VideoScene[]}, \texttt{DialogueSegment[]}, and
\texttt{ModalityVariants} types are provider-agnostic interfaces; any
modality backend can be swapped without touching the delivery layer.

\paragraph{Video backend upgrade path.}
The current in-browser renderer delivers animated typographic video at
zero infrastructure cost. Organisations requiring higher production
value can connect any of the following backends to the same
\texttt{VideoScene[]} interface: \textbf{Higgsfield}
\cite{higgsfield2026}, a multi-model creative platform aggregating
Sora~2, Kling~2.6, and Google Veo~3.1 with cinema-grade camera
simulation; \textbf{Wan~2.1} \cite{wan2025} (Alibaba, MIT
open-source), a text-to-video suite from 1.3B to 14B parameters
enabling 480p generation on a consumer GPU; and
\textbf{HunyuanVideo} \cite{hunyuan2025} (Tencent, open-source), a
Diffusion Transformer model suited for instructor avatar generation in
corporate training contexts.

The principle is: Sudar is the authoring and delivery operating system;
the generation backend is a plugin. A school in Sub-Saharan Africa
self-hosts Wan~2.1 on a local GPU server at zero per-call cost. A
Fortune~500 L\&D team connects Higgsfield for cinematic course
introductions. The learner experience---the Digital Learner Twin,
adaptive sequencing, AI tutor, and modality switching---is identical.

\paragraph{Community extension patterns.}
The MIT licence enables: new modality plugins (VR, AR, 3D simulation);
new LLM backends (any OpenAI-compatible API endpoint via environment
variable); new language and voice packs (Sarvam AI Indian English
neural voices are already integrated alongside Edge-TTS); and
vertical-specific forks for medical education, software engineering
bootcamps, and regulated-industry compliance.

\subsection{The Evolution of Sudar Intelligence}
\label{sec:intelligence}

The Intelligence layer described above reflects the current
implementation: reactive Q\&A via RAG, rule-based proactive nudges,
and a longitudinal memory updated from quiz outcomes and tutor
exchanges. This is a substantial advance over stateless LMS AI; it is
also, architecturally, only the beginning.

As the \texttt{learning\_events} table accumulates interaction data,
the Intelligence engine transitions from rule-based heuristics to
\emph{learned} intervention policies. The research frontier demonstrates
what becomes possible: reinforcement learning-driven ITS achieves 28.6\%
better intervention adaptability, 31.2\% reduction in recurring errors,
and 24.8\% lower dropout rates compared to rule-based systems
\cite{meng2025rltutor}.

Sudar's architecture anticipates this transition explicitly, with the
Digital Learner Twin and \texttt{learning\_events} stream serving as
the shared substrate for future first-party and community-built
Intelligence plugins: \textbf{SudarCoach} (career pathway advisor),
\textbf{SudarAssess} (item response theory psychometric layer),
\textbf{SudarSocial} (peer learning orchestrator),
\textbf{SudarCompliance} (automated evidence collection for regulated
industries), and \textbf{SudarTranslate} (real-time multilingual
delivery).

% ---------------------------------------------------------------
\section{Implementation and Reproducibility}
\label{sec:impl}
% ---------------------------------------------------------------

The reference platform is implemented with Next.js~14 (Studio, Learn;
TypeScript strict mode, App Router, Tailwind CSS),
Python~3.11+ FastAPI (Intelligence; async handlers, Pydantic v2), and
Supabase (PostgreSQL + pgvector, auth, storage). AI providers include
Together AI (primary; open-weight models), OpenAI, and Anthropic.
TTS uses Edge-TTS (default; zero cost, 40+ languages, 300+ voices)
with optional Sarvam AI Indian English neural voices. The schema and
event model are fully documented in the repository
(\texttt{ECOSYSTEM.md}, \texttt{RESEARCH\_FOUNDATION.md}).

Currently implemented: shared authentication and schema; course and
learning path CRUD and publish; document-to-course AI generation;
learner dashboard with streak, time-on-task, Sudar recommendations,
and upcoming deadlines; six modalities (text with sentence-level
read-along TTS, standalone Listen (audio TTS) tab, animated video,
audio podcast, interactive mindmap, flashcards, SCORM delivery); AI
tutor with longitudinal memory, RAG, text-selection popup, resizable
overlay, ``Your context'' live memory panel, and generative block
rendering; five quiz mode archetypes and five lesson archetypes;
next-best-action and adaptive path ordering; learning paths with
unlock rules and shareable certification links with PDF generation;
compliance view (overdue / at-risk / on-track), compliance email
reminders (cron for at-risk/overdue path assignments); analytics
(completions, quiz scores, struggle topics, time per section); and
learner preferences page. Production deployment is documented:
Studio and Learn host on Vercel; Intelligence hosts on Railway,
Render, or Fly.io, with the Intelligence URL configured in both
front-end projects. The project is open source under the MIT
licence \cite{sudar2026}.

% ---------------------------------------------------------------
\section{Economic Accessibility: The Radical Cost Case}
\label{sec:economics}
% ---------------------------------------------------------------

The economic dimension of Sudar is not incidental to its design---it
is one of its three primary contributions.

\paragraph{Observed infrastructure cost.}
Operating Sudar at full engagement---10 courses generated, 20 learner
review sessions with AI tutor chat, adaptive personalisation, mindmap
generation, flashcard generation, and TTS narration---costs
approximately \textbf{\$0.20 per week} on the Together AI + Edge-TTS
stack. At organisational scale with 1,000 learners and 50 sessions per
month, the contrast is:

\begin{center}
\small
\begin{tabular}{@{}lrr@{}}
\toprule
Stack & Per month & Annual \\
\midrule
Sudar (Together AI + Edge-TTS) & \$21 & \$252 \\
GPT-4o + OpenAI TTS-1 & \$3,405 & \$40,860 \\
Claude 3.5 Sonnet + Azure TTS & \$3,735 & \$44,820 \\
Docebo (platform licence) & \$5,830+ & \$69,960+ \\
\bottomrule
\end{tabular}
\end{center}

At 10,000 learners, Sudar's open-weight stack costs approximately
\$210/month. The delta versus a closed-model stack---over \$400,000
per year---is the economic value of open-weight inference returned to
the institution rather than captured by AI API providers.

\paragraph{The inference cost collapse.}
AI inference costs have fallen approximately 1,000-fold from 2022 to
2026 at roughly 10$\times$ per year \cite{a16z2024llmflation}, driven
by GPU efficiency gains, software optimisation (vLLM, TensorRT-LLM),
Mixture-of-Experts architectures, and quantisation. GPT-3-class
performance costs \$0.06 per million tokens today versus \$60 in 2021.
The cost of delivering a world-class personalised tutor interaction to
one learner is now less than \$0.02 per month---below the cost of a
single SMS message. The economic argument against AI-native learning at
scale has expired.

\paragraph{The education access crisis.}
As of 2024, 251 million children and youth remain out of school
globally \cite{unesco2024}. 2.6 billion people remain offline
\cite{itu2024}, with mobile internet in Africa costing 14$\times$ more
than in Europe. An annual financing gap of \$39 billion exists for
basic education in low- and lower-middle-income countries through 2030
\cite{unesco2024}. The average corporate training spend per employee in
the United States is \$1,283 per year \cite{statista2023}; Sudar's
infrastructure cost of \$0.21/user/month (\$2.52/year) represents a
99.8\% cost reduction for the AI infrastructure layer relative to
this benchmark.

High-end AI learning platforms priced at \$12--\$20 per user per month
are not deployable in this context. A platform whose full AI
infrastructure costs \$0.02 per learner per month---and whose stack is
open-source, self-hostable, and runnable on open-weight models---is.

% ---------------------------------------------------------------
\section{Evidence and Evaluation Strategy}
\label{sec:eval}
% ---------------------------------------------------------------

\paragraph{Research foundation.}
The design is evidence-informed. \texttt{RESEARCH\_FOUNDATION.md}
maps each capability to learning-science references: adaptive
instruction \cite{vanlehn2011,aleven2016}, dual coding
\cite{mayer2009,clark2016}, self-regulated learning
\cite{zimmerman2002}, formative assessment \cite{black1998,roediger2006},
intelligent tutoring \cite{self1988,vanlehn2011}, and learner
modelling \cite{bull2016,self1988}.

\paragraph{Scope of current claims.}
This is a system and architecture paper. Current claims are:
(a)~the platform implements all described components to the level of a
working proof of concept; (b)~each component is grounded in
established learning-science evidence; (c)~the ALP architecture
addresses a gap that no widely adopted technology currently fills; and
(d)~the economic cost structure has been empirically observed in
practice and validated against current provider pricing.

\paragraph{Pilot study.}
A pilot evaluation is planned, to be initiated after the current build
phase, and will be conducted in partnership with an organisational
partner to assess adoption, engagement, and early learning outcomes.
Results will be reported in a subsequent paper when available.

\paragraph{Future controlled studies.}
Plans include randomised or quasi-experimental comparisons of Sudar
versus a standard LMS on learning outcomes, learner satisfaction, and
time-to-competency, as well as studies isolating the contribution of
the longitudinal tutor memory component and the cost-per-outcome metric
versus incumbent platforms.

% ---------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}
% ---------------------------------------------------------------

\paragraph{Limitations.}
Production ALP plugin connectors for Moodle, Canvas, and Blackboard
are not yet built; the API surface is defined but LMS-specific adapter
layers require additional engineering. Controlled efficacy data are not
yet available; claims about learning gains remain grounded in the
underlying learning-science evidence rather than Sudar-specific trials.
The Digital Learner Twin relies on inferred signals that may not fully
capture a learner's actual knowledge state \cite{self1988,bull2016}.
Edge-TTS accesses Microsoft's Azure Neural Voice infrastructure without
a formal commercial licence; production deployments with paying
customers should provision Azure AI Speech or an equivalent licensed
TTS service. The predictive intelligence capabilities described in
Section~\ref{sec:intelligence} require longitudinal event data to train
effectively; early deployments will rely on rule-based proactive nudges
until sufficient behavioural data accumulates.

\paragraph{Future work.}
Planned work includes: LMS-specific connectors for ALP (Moodle, Canvas,
Blackboard); completing SudarPlay (gamified modality) and SudarFeed
(social feed modality); developing the predictive intervention layer
with RL-based policy learning; running controlled efficacy studies;
releasing SudarCoach, SudarAssess, and SudarCompliance as first-party
Intelligence plugins; building the community plugin ecosystem; and
supporting enterprise features (SSO, HRIS integration) and an
offline-first variant for connectivity-constrained environments.

% ---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
% ---------------------------------------------------------------

We presented Sudar, an AI-native learning platform with three
contributions: (1)~a reference platform combining authoring, delivery,
and intelligence around a Digital Learner Twin, adaptive sequencing, a
complete multimodal delivery stack, and an AI tutor with longitudinal
cross-session memory; (2)~the Adaptive Learning Layer (ALP), an
architecture for deploying these capabilities as a plugin layer on top
of existing LMSs---with the potential to reach the hundreds of millions
of Moodle, Canvas, Blackboard, and Sakai users \cite{moodle2025stats}
without requiring LMS replacement; and (3)~a demonstrated economic
model that reduces the infrastructure cost of AI-native personalised
learning to less than \$0.02 per learner per month, making it viable
at the scale of national education systems and accessible in contexts
where \$12--\$20/user/month platforms are structurally excluded.

The inference cost collapse is the macro condition that makes all of
this possible \cite{a16z2024llmflation}. Platforms built on open-weight
models and open-source infrastructure are positioned to pass this
economic gain directly to learners---particularly the 251 million
children currently out of school \cite{unesco2024} and the 2.6 billion
people currently offline \cite{itu2024}.

Together, these contributions offer a path to making adaptive,
memory-aware learning accessible both through a standalone platform and
as an augmentation of the LMSs that organisations already rely on. The
reference implementation is open source, grounded in a broad
learning-science evidence base
\cite{vanlehn2011,aleven2016,mayer2009,clark2016,zimmerman2002,%
gagne1985,bull2016,black1998,roediger2006}, and designed as a bedrock
upon which the community can build modalities, intelligence plugins,
and LMS connectors.

% ---------------------------------------------------------------
\section*{Acknowledgments}
% ---------------------------------------------------------------

The system and research are the author's project. The reference
implementation is available at
\url{https://github.com/Dhanikesh-Karunanithi/Sudar} with
documentation (\texttt{ECOSYSTEM.md}, \texttt{RESEARCH\_FOUNDATION.md},
\texttt{docs/sudar-memory.md}).

\begin{center}
\fbox{\parbox{0.9\columnwidth}{\small
\textbf{Call for collaboration.} Institutions, organisations, and
open-source contributors are invited to collaborate on pilots, plugin
integrations, and community extensions.
Contact: \texttt{connect@dhanikeshkarunanithi.com} or
\url{https://github.com/Dhanikesh-Karunanithi/Sudar}.
}}
\end{center}

% ---------------------------------------------------------------
\appendix
\section{Schema Summary}
% ---------------------------------------------------------------

Key tables for reproducibility:
\texttt{learner\_profiles} (Digital Learner Twin:
\texttt{ai\_tutor\_context}, \texttt{next\_best\_action}, modality
affinity scores, behavioural signals, \texttt{preferences} JSONB);
\texttt{learning\_events} (\texttt{event\_type}, payload, modality,
duration);
\texttt{ai\_interactions} (\texttt{user\_message},
\texttt{ai\_response}, \texttt{context\_used});
\texttt{enrollments} (path/course, progress, \texttt{due\_date});
\texttt{certifications}.
Full schema: \texttt{ECOSYSTEM.md} in the repository.

\section{Infrastructure Cost Reference}

\begin{center}
\small
\begin{tabular}{@{}p{3.2cm}rr@{}}
\toprule
Stack & /learner/mo & Annual (1k) \\
\midrule
Together AI 8B + Edge-TTS   & \$0.021  & \$252 \\
Together AI 70B + Edge-TTS  & \$0.105  & \$1,260 \\
GPT-4o + OpenAI TTS-1       & \$3.41   & \$40,860 \\
Claude 3.5 + Azure TTS      & \$3.74   & \$44,820 \\
Docebo (licence, low end)   & \$5.83   & \$69,960 \\
Sana Labs (est., min.\ 300) & $\approx$\$15 & $\approx$\$180k \\
\bottomrule
\end{tabular}
\end{center}
\noindent\small{AI infrastructure only; excludes hosting (Supabase free
tier: \$0; Pro: \$25/month).}

% ---------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{references}
\end{document}
